One-dimensional turbulence (ODT): computationally efficient modeling and simulation of turbulent flows

### 2020-09-21 manuscript/manuscript.tex
conclusions section added, spell checked paper.

Turbulent flows characterize the vast majority of fluid flows in practical engineering applications, and simulations of turbulent flows provide researchers with valuable insights into complex systems, particularly reacting turbulent flows such as combustion processes. Turbulence is a complex phemonenon that affects the full range of a flow's length and time scales. As a result, resolving the entire flow field by numerically solving the Navier-Stokes equations of fluid flow, as is done in direct numerical simulations (DNS), requires substantial computational resources. DNS is a powerful research tool, but its high computational cost makes it intractable for simulating most practical engineering flows. In order to achieve numerical solutions to practical flow problems, researchers can use alternative frameworks that model turbulence rather than resolving it directly.

@@

Early applications of ODT focused on homogenous turbulence, wakes, and mixing layers \cite{Kerstein_1999,Kerstein_2000,Kerstein_2001}. Later extension to variable-density flows and a spatial downstream coordinate system facilitated its growth and application to more complex flows, including combustion in jet flames \cite{Echekki_2001,Hewson_2001,Hewson_2002,Lignell_2012,Punati_2011,Abdelsamie_2017,Lignell_2017, Goshayeshi_2015}, counterflow flames \cite{Jozefik_2015}, wall fires \cite{Monson_2016}, and sooting flames \cite{Lignell_2015,Hewson_2006,Hewson_2009,Lignell_2015b,Ricks_2010}, as well as other particle flows \cite{Sun_2017,Schmidt_2009,Sun_2014,Fistler_2017}. ODT has also served to complement LES through subgrid modeling studies \cite{Cao_2008,Schmidt_2003,Schmidt_2010} and has been applied to various other flow configurations such as double-diffusive interfaces \cite{GonzalezJuez_2011}, Rayleigh-Taylor mixing \cite{GonzalezJuez_2013}, and stratified turbulence \cite{Wunsch_2001}. Most recently, the ODT code was extended to include cylindrical and spherical coordinate systems \cite{Lignell_2018,Klein_2018,Klein_2019}.

@@

Eddy events modify domain variables using ``triplet maps", as illustrated in Figure~\ref{fig:tripletmap}. In an eddy region of size $l$, transported property profiles are modified (mapped) by making three copies, compressing each spatially by a factor of three, then lining these up with the middle copy spatially reversed.
This map is continuous and conservers all quantities and statistical moments, increases scalar gradients and decreases length scales consistent with turbulent eddies in real flows. Subsequent eddies in the same region will result in a cascade of scales. The eddy rates, used in the sampling and selection of eddies depend on eddy size and the local kinetic energy consistent with turbulent scaling laws.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{../figures/tripletmap/tripmap.pdf} 
    \caption{Schematic diagram of a triplet map. From left to right: (first pane) simulation of a single ``eddy", a two-dimensional Kelvin-Helmholz instability with a vertical white line-of-sight; (second pane) mixture fraction profile along the indicated line-of-sight before and after the eddy showing increased gradients and surface area; (third pane) a triplet map operates on a profile to produce a mapped profile with three copies, compressed and lined up, with the middle copy spatially inverted; (fourth pane) illustrative eddy event sizes and locations with corresponding temperature profile for a single ODT realization of a wall fire, adapted from \cite{Monson_2016}.}
	\label{fig:tripletmap}
\end{figure}

@@

The ODT build process is automated with CMake. The user navigates to the \texttt{build} directory and edits the CMake configuration file. The CMake configuration file specifies Cantera's location and must be changed to reflect the local installation location. For reacting flows, the chemical reation rates can be computed by Cantera (the default), or a from a user-specified function, specified at compile time. Once the configuration settings are updated, the user runs CMake to apply the changes.

@@

ODT simulations can also be run in parallel using MPI. Two run scripts, \texttt{slrmJob.sh} and \texttt{slrmJob\textunderscore array.sh}, are configured using SLURM \cite{Yoo_2003}, a common workload manager used for massively parallel computing resources. This allows many ODT realizations to run in parallel rather than in serial, reducing overall simulation time. Individual realizations are independent (\emph{embarrassingly parallel}), but users must take care with case names and input file changes to ensure that individual realizations or entire cases are not overridden accidentally.

@@

The ODT model formulation is somewhat complex and is difficult to cleanly separate from its implementation. For instance, while the triplet map can be represented mathmatically, it is usually implemented as we described previously. That description assumed an adaptive grid with variable cell sizes. Alternative formulations are used for fixed grids. Furthermore, the model itself is complex and difficult to implement. These challenges present barriers for practitioners who wish to use the model. The code presented aims to reduce these barriers by providing an open-source, well-documented code that can be used directly, or modified to suit particular research aims. New research questions that could be persued include soot-flame-radiation-chemsitry interactions, multi-phase reacting flows (such as diesel sprays) using Lagrangian particles, and coupled near-wall heat and momentum transfer, to name a few. The code presented has been demonstrated to be extensible to such systems. These extensions are facilitated by the object-oriented structure, including the domain variables. For example, the code has be extended to solve steady laminar flamelets \cite{Peters_1984}, requiring inheritance from the \texttt{solver}, \texttt{micromixer}, and \texttt{dv} (domain variable) classes.

The structure of the ODT code, using an adaptive mesh, and a Lagrangian formulation for diffusive advancement of transport equations has several advantages. These include simplicity of formulation, and computational efficiency. The adaptive mesh minimizes the number of computational cells required, and can be adapted on any variable of interested. (Older versions of the code even made use of multiple coupled computational grids for different scalars.) This is especially important for flows with complex and expensive chemistry, and flows with widely varying resolution demands (such as near-wall flows). The Lagrangian formulation, implies that cell faces move with fluid velocity (such as is present during, e.g., flame expansion). This treatment is inherently stable, removing the need for complex (and diffusive) hyberbolic conservation law treatements and their associated CFL constraints. The conservation equations, as in Eq. (\ref{eq:transport}), are similarly simplified.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
\label{acknoledgements}

This work was supported in part by the National Science Foundation under Grant No. CBET-1403403.

>>>>

Turbulent flows characterize the vast majority of fluid flows in practical engineering applications, and simulations of turbulent flows provide researchers with valuable insights into complex systems, particularly reacting turbulent flows such as combustion processes. Turbulence is a complex phenomenon that affects the full range of a flow's length and time scales. As a result, resolving the entire flow field by numerically solving the Navier-Stokes equations of fluid flow, as is done in direct numerical simulations (DNS), requires substantial computational resources. DNS is a powerful research tool, but its high computational cost makes it intractable for simulating most practical engineering flows. In order to achieve numerical solutions to practical flow problems, researchers can use alternative frameworks that model turbulence rather than resolving it directly.

@@

Early applications of ODT focused on homogeneous turbulence, wakes, and mixing layers \cite{Kerstein_1999,Kerstein_2000,Kerstein_2001}. Later extension to variable-density flows and a spatial downstream coordinate system facilitated its growth and application to more complex flows, including combustion in jet flames \cite{Echekki_2001,Hewson_2001,Hewson_2002,Lignell_2012,Punati_2011,Abdelsamie_2017,Lignell_2017, Goshayeshi_2015}, counterflow flames \cite{Jozefik_2015}, wall fires \cite{Monson_2016}, and sooting flames \cite{Lignell_2015,Hewson_2006,Hewson_2009,Lignell_2015b,Ricks_2010}, as well as other particle flows \cite{Sun_2017,Schmidt_2009,Sun_2014,Fistler_2017}. ODT has also served to complement LES through subgrid modeling studies \cite{Cao_2008,Schmidt_2003,Schmidt_2010} and has been applied to various other flow configurations such as double-diffusive interfaces \cite{GonzalezJuez_2011}, Rayleigh-Taylor mixing \cite{GonzalezJuez_2013}, and stratified turbulence \cite{Wunsch_2001}. Most recently, the ODT code was extended to include cylindrical and spherical coordinate systems \cite{Lignell_2018,Klein_2018,Klein_2019}.

@@

Eddy events modify domain variables using ``triplet maps", as illustrated in Figure~\ref{fig:tripletmap}. In an eddy region of size $l$, transported property profiles are modified (mapped) by making three copies, compressing each spatially by a factor of three, then lining these up with the middle copy spatially reversed.
This map is continuous and conserves all quantities and statistical moments, increases scalar gradients and decreases length scales consistent with turbulent eddies in real flows. Subsequent eddies in the same region will result in a cascade of scales. The eddy rates, used in the sampling and selection of eddies depend on eddy size and the local kinetic energy consistent with turbulent scaling laws.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{../figures/tripletmap/tripmap.pdf} 
    \caption{Schematic diagram of a triplet map. From left to right: (first pane) simulation of a single ``eddy", a two-dimensional Kelvin-Helmholtz instability with a vertical white line-of-sight; (second pane) mixture fraction profile along the indicated line-of-sight before and after the eddy showing increased gradients and surface area; (third pane) a triplet map operates on a profile to produce a mapped profile with three copies, compressed and lined up, with the middle copy spatially inverted; (fourth pane) illustrative eddy event sizes and locations with corresponding temperature profile for a single ODT realization of a wall fire, adapted from \cite{Monson_2016}.}
	\label{fig:tripletmap}
\end{figure}

@@

The ODT build process is automated with CMake. The user navigates to the \texttt{build} directory and edits the CMake configuration file. The CMake configuration file specifies Cantera's location and must be changed to reflect the local installation location. For reacting flows, the chemical reaction rates can be computed by Cantera (the default), or a from a user-specified function, specified at compile time. Once the configuration settings are updated, the user runs CMake to apply the changes.

@@

ODT simulations can also be run in parallel using MPI. Two run scripts, \texttt{slrmJob.sh} and \texttt{slrmJob\textunderscore array.sh}, are configured using Slurm \cite{Yoo_2003}, a common workload manager used for massively parallel computing resources. This allows many ODT realizations to run in parallel rather than in serial, reducing overall simulation time. Individual realizations are independent (\emph{embarrassingly parallel}), but users must take care with case names and input file changes to ensure that individual realizations or entire cases are not overridden accidentally.

@@

The ODT model formulation is somewhat complex and is difficult to cleanly separate from its implementation. For instance, while the triplet map can be represented mathematically, it is usually implemented as we described previously. That description assumed an adaptive grid with variable cell sizes. Alternative formulations are used for fixed grids. Furthermore, the model itself is complex and difficult to implement. These challenges present barriers for practitioners who wish to use the model. The code presented aims to reduce these barriers by providing an open-source, well-documented code that can be used directly, or modified to suit particular research aims. New research questions that could be pursued include soot-flame-radiation-chemistry interactions, multi-phase reacting flows (such as diesel sprays) using Lagrangian particles, and coupled near-wall heat and momentum transfer, to name a few. The code presented has been demonstrated to be extensible to such systems. These extensions are facilitated by the object-oriented structure, including the domain variables. For example, the code has be extended to solve steady laminar flamelets \cite{Peters_1984}, requiring inheritance from the \texttt{solver}, \texttt{micromixer}, and \texttt{dv} (domain variable) classes.

The structure of the ODT code, using an adaptive mesh, and a Lagrangian formulation for diffusive advancement of transport equations has several advantages. These include simplicity of formulation, and computational efficiency. The adaptive mesh minimizes the number of computational cells required, and can be adapted on any variable of interested. (Older versions of the code even made use of multiple coupled computational grids for different scalars.) This is especially important for flows with complex and expensive chemistry, and flows with widely varying resolution demands (such as near-wall flows). The Lagrangian formulation, implies that cell faces move with fluid velocity (such as is present during, e.g., flame expansion). This treatment is inherently stable, removing the need for complex (and diffusive) hyperbolic conservation law treatments and their associated CFL constraints. The conservation equations, as in Eq. (\ref{eq:transport}), are similarly simplified.

@@

An implementation of the One-Dimensional Turbulence model has been presented. The code is open-source and uses is based on an object-oriented C++ framework. The code was motivated as an advanced model for turbulent reacting flow research. The build process (using CMake), was described, along with the directory structure, input file format, output data, and run and post-processing. A reacting jet example was given along with comparison to experimental data. 
Access to a proven, documented code, and a modern build procedure will aid researchers in applying and extending the model.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
\label{acknowledgements}

This work was supported in part by the National Science Foundation under Grant No. CBET-1403403.
%TODO, we probably need to add some acknowlegements here for those who contributed to the code.
